### 2021-11-04 Rmd/00-e-preface.Rmd

# Preface {-}

Chapter \@ref(ch:gravitas) has been accepted by the *Journal of Computational and Graphical Statistics*. It has won the ACEMS Business Analytics Prize in 2020. The accompanying R package **gravitas** is on CRAN. Chapter \@ref(ch:hakear) and Chapter \@ref(ch:gracsr) are yet to be submitted.

### Open and reproducible research {-}

This thesis is written in R Markdown [@rmarkdown] with **bookdown** [@bookdown], using **renv** [@renv] to create reproducible environments. The online version of this thesis is hosted at <https://sayani.netlify.app/>, powered by [Netlify](http://netlify.com). All materials (including the data sets and source files) required to reproduce this document can be found at the Github repository <https://github.com/Sayani07/thesis>.

### License {-}

<a rel="license" href="http://creativecommons.org/licenses/by-nc-sa/4.0/"><img alt="Creative Commons Licence" style="border-width:0" src="https://i.creativecommons.org/l/by-nc-sa/4.0/88x31.png" /></a>

This work is licensed under a [Creative Commons Attribution-NonCommercial-ShareAlike 4.0 International License](http://creativecommons.org/licenses/by-nc-sa/4.0/).

The code used in this document is available under the [MIT license](https://opensource.org/licenses/MIT).

\clearpage\pagenumbering{arabic}\setcounter{page}{0}