### 2021-11-04 Rmd/04-gracsr.Rmd

# Data representation, visual and analytical techniques for demystifying temporal missing data {#ch:mists}