### 2021-11-04 Rmd/00-d-acknowlegements.Rmd
remove last 2 commits

# Acknowledgements {-}

## supervisors
Rob and Di are phenomenal leaders in their respective fields, who lead by example. I am inspired by their creativity, wisdom, discipline, and dedication to really contribute to the society through their research. Thank you for constantly pushing me to improve as a programmer and researcher, and for regularly sharing best practices for conducting research. Looking back, I am grateful for how my thoughts and work in statistical computing, graphics, and data analytics in general have evolved over the years. I still have a lot to learn, but I am a much more self-reliant and independent researcher than I was when I started. Di has been instrumental in exposing me to the potential benefits of effective data visualization. As a female researcher, I am encouraged by her willingness to pursue unconventional avenues and have frequently noticed her make a conscious decision to question existing  stereotypes and biases. Thank you Di for being a fantastic female role model.

## department and PhD colleagues

## R community

## friends

PhD in Covid has been harder and I would not have been able to make it through this if not for the emotional support of my friends and family. Thank you Puwasala, Sium for being an epitome of kindness and acting as a pillar of support. Thank you Tushar, Samarpita and Nairita for always motivating me and believing in me when I didn't believe in myself. Thank you Ian for always putting things in perspective when I lost sight. Thank you each one of you for being available, listening to me, and not giving up on me at my lowest. Thanks to my housemates Anjali, Surbhi and Dulaji for the fun company and conversations that kept me sane in the covid lockdowns.

## family
Finally, a big thanks to my family for always being supportive of my choices. Thanks to my mum (Nupur Gupta) and dad (Arun Prasad Gupta), from whom I learnt that no matter where you start from, if you persevere and are sincere in your efforts, you can sail through any difficult situation. They are my constant cheerleaders for all little and big endeavours. Thanks to my brother (Avijit) for always having my back and inspiring me to dream bigger. Thank you, Juhi (sister-in-law) for instilling in me the value of organisation in all aspects of life and the importance of prioritizing my physical health from time to time. Thank you all for your part in supporting me in this journey. I love you.