### 2021-11-04 thesis.Rmd

---
title: "Visualization and analysis of probability distributions of large temporal data"
author: "Sayani Gupta"
degreetype: Doctor of Philosophy
site: bookdown::bookdown_site
degrees: M.Stat, B.Sc (Stat Hons)
bibliography:
- bib/gravitas.bib
- bib/hakear.bib
- bib/gracsr.bib
- bib/thesisrefs.bib
link-citations: yes
---

```{r initial, echo = FALSE, cache = FALSE, results = 'hide'}
options("knitr.graphics.auto_pdf" = TRUE)
library(knitr)
opts_chunk$set(
  warning = FALSE, message = FALSE, echo = FALSE,
  fig.path = 'figure/', fig.align = 'center', fig.show = 'hold',
  cache = TRUE, external = TRUE, comment = "#>", dpi = 300,
  out.width = ifelse(is_html_output(), "100%", "\\textwidth"),
  cache.path = ifelse(is_html_output(), 'cache-gitbook/', 'cache-pdfbook/')
)
hook_output <-  knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})
```

`r if (knitr::is_latex_output()) '<!--'`

# Welcome {-}

This is the website for my PhD thesis at Monash University (Australia), titled "Visualization and analysis of probability distributions of large temporal data".

`r Sys.Date()`

`r if (knitr::is_latex_output()) '-->'`

<!--chapter:end:index.Rmd-->

`r if (knitr::is_html_output()) '<!--'`

# Copyright notice {-}

\textcopyright { }Sayani Gupta (\number\the\year).

I certify that I have made all reasonable efforts to secure copyright permissions for third-party content included in this thesis and have not knowingly added copyright content to my work without the owner's permission.

\newpage

`r if (knitr::is_html_output()) '-->'`

<!--chapter:end:Rmd//00-a-copyright.Rmd-->

# Abstract {-}

The research was motivated by the desire to understand the Australian smart meters data, which was collected half-hourly for two years at the household level. With temporal data available at ever finer scales, exploring periodicity can become overwhelming with so many possible temporal deconstructions to explore. Analysts are expected to comprehensively explore the many ways to view and consider temporal data. However, the plethora of choices and the lack of a systematic approach to do so quickly can make the task overwhelming.

<!-- exploration may be required to be carried out across both finer and coarser scales to draw useful inferences about the underlying process. It is important to be able to navigate through all the temporal deconstructions that accommodate for periodicities to have multiple perspectives of the observed data. This idea aligns with the notion of EDA (Tukey 1977) which emphasizes the use of multiple perspectives on data to help formulate hypotheses before proceeding to hypothesis testing. Slicing and dicing the data in all possible temporal scales as suggested by EDA can be a daunting task as it leads to a -->
<!-- myriad of possibilities. -->

<!-- Analysts are expected to comprehensively explore the many ways to view and consider temporal data. However, the plethora of choices and the lack of a systematic approach to do so quickly can make the task overwhelming -->

<!-- exploring periodicity can become overwhelming with so many possible temporal deconstructions to explore. Analysts are expected to comprehensively explore the many ways to view and consider temporal data. However, the plethora of choices and the lack of a systematic approach to do so quickly can make the task overwhelming. -->

<!-- <WHY> <WHAT> <HOW> -->

This work investigates how time may be dissected, resulting in alternative data segmentation and, as a result, different visualizations that can aid in the identification of underlying patterns. The first contribution (Chapter \@ref(ch:gravitas)) describes classes of time deconstructions using linear and cyclic time granularities. It provides tools to compute possible cyclic granularities from an ordered (usually temporal) index and also a framework to systematically explore the distribution of a univariate variable conditional on two cyclic time granularities by defining "harmony". "harmony" denotes pairs of granularities that could be analyzed together and reduces the search from all possible options. This approach is still overwhelming for human consumption because there would still be a huge number of harmonies. Hence, the second contribution (Chapter \@ref(ch:hakear)) refines the search of informative granularities by identifying those for which the differences between the displayed distributions are greatest and rating them in order of importance of capturing maximum variation. The third contribution (Chapter \@ref(ch:gracsr)) builds upon the first two to provide methods for exploring heterogeneities in repetitive behavior for many households and over multiple granularities. It accomplishes this by providing a way to cluster time series based on probability distributions across informative cyclic granularities. Although we were motivated by the smart meter example, the problem and the solutions we propose are practically relevant to any temporal data observed more than once per year.

<!-- there is a need to resort to clustering, which has been addressed in  It approaches the problem by computing pairwise distances between time series based on probability distributions over multiple cyclic granularities at once. -->
\newpage

<!--chapter:end:Rmd//00-b-abstract.Rmd-->

`r if (knitr::is_html_output()) '<!--'`

\newpage

# Declaration {-}

I have not renumbered sections of submitted or published papers in order to generate a consistent presentation within the thesis.

\textbf{Student name:} Sayani Gupta

\textbf{Student signature:}

```{r signature, out.width = "20%", fig.align = "left", eval = file.exists("img/signature.png")}
knitr::include_graphics("img/signature.png")
```

\textbf{Date:} `r Sys.Date()`

`r if (knitr::is_html_output()) '-->'`

<!--chapter:end:Rmd//00-c-declaration.Rmd-->

# Acknowledgements {-}

## supervisors
Rob and Di are phenomenal leaders in their respective fields, who lead by example. I am inspired by their creativity, wisdom, discipline, and dedication to really contribute to the society through their research. Thank you for constantly pushing me to improve as a programmer and researcher, and for regularly sharing best practices for conducting research. Looking back, I am grateful for how my thoughts and work in statistical computing, graphics, and data analytics in general have evolved over the years. I still have a lot to learn, but I am a much more self-reliant and independent researcher than I was when I started. Di has been instrumental in exposing me to the potential benefits of effective data visualization. As a female researcher, I am encouraged by her willingness to pursue unconventional avenues and have frequently noticed her make a conscious decision to question existing  stereotypes and biases. Thank you Di for being a fantastic female role model.

## department and PhD colleagues

## R community

## friends

PhD in Covid has been harder and I would not have been able to make it through this if not for the emotional support of my friends and family. Thank you Puwasala, Sium for being an epitome of kindness and acting as a pillar of support. Thank you Tushar, Samarpita and Nairita for always motivating me and believing in me when I didn't believe in myself. Thank you Ian for always putting things in perspective when I lost sight. Thank you each one of you for being available, listening to me, and not giving up on me at my lowest. Thanks to my housemates Anjali, Surbhi and Dulaji for the fun company and conversations that kept me sane in the covid lockdowns.

## family
Finally, a big thanks to my family for always being supportive of my choices. Thanks to my mum (Nupur Gupta) and dad (Arun Prasad Gupta), from whom I learnt that no matter where you start from, if you persevere and are sincere in your efforts, you can sail through any difficult situation. They are my constant cheerleaders for all little and big endeavours. Thanks to my brother (Avijit) for always having my back and inspiring me to dream bigger. Thank you, Juhi (sister-in-law) for instilling in me the value of organisation in all aspects of life and the importance of prioritizing my physical health from time to time. Thank you all for your part in supporting me in this journey. I love you.

<!--chapter:end:Rmd//00-d-acknowlegements.Rmd-->

# Preface {-}

Chapter \@ref(ch:gravitas) has been accepted by the *Journal of Computational and Graphical Statistics*. It has won the ACEMS Business Analytics Prize in 2020. The accompanying R package **gravitas** is on CRAN. Chapter \@ref(ch:hakear) and Chapter \@ref(ch:gracsr) are yet to be submitted.

### Open and reproducible research {-}

This thesis is written in R Markdown [@rmarkdown] with **bookdown** [@bookdown], using **renv** [@renv] to create reproducible environments. The online version of this thesis is hosted at <https://sayani.netlify.app/>, powered by [Netlify](http://netlify.com). All materials (including the data sets and source files) required to reproduce this document can be found at the Github repository <https://github.com/Sayani07/thesis>.

### License {-}

<a rel="license" href="http://creativecommons.org/licenses/by-nc-sa/4.0/"><img alt="Creative Commons Licence" style="border-width:0" src="https://i.creativecommons.org/l/by-nc-sa/4.0/88x31.png" /></a>

This work is licensed under a [Creative Commons Attribution-NonCommercial-ShareAlike 4.0 International License](http://creativecommons.org/licenses/by-nc-sa/4.0/).

The code used in this document is available under the [MIT license](https://opensource.org/licenses/MIT).

\clearpage\pagenumbering{arabic}\setcounter{page}{0}

<!--chapter:end:Rmd//00-e-preface.Rmd-->

# Introduction {#ch:intro}

Placeholder

## Visualizing probability distributions across bivariate cyclic temporal granularities {#sec:gravitas}
## Detecting distributional differences between temporal granularities for exploratory time series analysis {#sec:hakear}
## Clustering time series based on probability distributions across temporal granularities {#sec:gracsr}
## Summary

<!--chapter:end:Rmd//01-intro.Rmd-->

# Visualizing probability distributions across bivariate cyclic temporal granularities {#ch:gravitas}

Placeholder

## Introduction
## Linear time granularities {#sec:linear-time}
### Definitions
### Relativities
## Cyclic time granularities {#sec:cyclic-gran}
### Circular granularities {#sec:circular-gran-def}
### Quasi-circular granularities {#sec:quasi-circular-gran-def}
### Aperiodic granularities {#sec:aperiodic-gran-def}
### Relativities {#sec:cyclic-calendar}
### Computation
#### Single-to-multiple order-up {-}
#### Multiple-to-single order-up {-}
#### Multiple order-up quasi-circular granularities {-}
## Data structure {#sec:data-structure}
### Harmonies and clashes {#sec:synergy}
### Near-clashes {#sec:near-clashes}
## Visualization {#sec:visualization}
### Data summarization
### Display choices for univariate distributions
### Comparison across sub-groups induced by conditioning {#sec:aesthetics}
#### Levels {-}
#### Synergy of cyclic granularities {-}
#### Conditioning variable {-}
## Applications {#sec:application}
### Smart meter data of Australia {#sec:smartmeter}
#### Cyclic granularities search and computation {-}
#### Screening and visualizing harmonies {-}
### T20 cricket data of Indian Premier League {#sec:cricket}
## Discussion {#sec:discussion}
## Acknowledgments {-}
## Supplementary Materials

<!--chapter:end:Rmd//02-gravitas.Rmd-->

# Detecting distributional differences between temporal granularities for exploratory time series analysis  {#ch:hakear}

Placeholder

## Introduction
## Proposed distance measure {#sec:computation-wpd}
### Principle 
### Notations
### Computation
#### Data transformation
#### Characterising distributions
#### Distance between distributions
#### Within-facet and between-facet distances
#### Tuning parameter
#### Raw distance measure
#### Adjusting for the number of comparisons {#sec:norm-wpd}
## Ranking and selection of cyclic granularities {#sec:rank-wpd}
### Selection
### Ranking
## Simulations {#sec:simulations}
### Behavior of raw and adjusted distance measures
### Choosing threshold
### Environment
## Application to residential smart meter dataset {#sec:application-wpd}
## Discussion
## Acknowledgments {-}
## Supplementary Materials

<!--chapter:end:Rmd//03-hakear.Rmd-->

# Clustering time series based on probability distributions across temporal granularities {#ch:gracsr}

Placeholder

## Introduction
## Clustering methodology {#sec:methodology}
## Validation {#sec:validation}
### Data generating processes
### Data designs  
#### Individual granularities
#### Interaction of granularities
### Visual exploration of findings
## Case study {#sec:cases}
### Clustering
#### JS-based distances
#### wpd-based distances
## Conclusion and future work {#sec:conclusion}
## Acknowledgments {-}
## Supplementary Materials

<!--chapter:end:Rmd//04-gracsr.Rmd-->

# Conclusion and future plans {#ch:conclusion}

Placeholder

## Software development
### gravitas
### hakear 
### gracsr
## Future work
### Putting all functionalities on CRAN
### Scaling up the clustering method to incorporate large uncertainty and improved computational efficiency that comes with large data
### Comparing our clustering method with other benchmark methods
### Check generalizability of our methods
## Final words

<!--chapter:end:Rmd//05-conclusion.Rmd-->

# (APPENDIX) Appendix {-}

# Data dictionary

```{r wdi-data-dict}
```

(ref:wdi-data-dict) Data dictionary for the World Development Indicators.

<!--chapter:end:Rmd//06-appendix.Rmd-->

`r if (knitr::is_html_output()) '# Bibliography {-}'`

<!--chapter:end:Rmd//99-references.Rmd-->