### 2021-11-04 index.Rmd

---
title: "Visualization and analysis of probability distributions of large temporal data"
author: "Sayani Gupta"
degreetype: Doctor of Philosophy
site: bookdown::bookdown_site
degrees: M.Stat, B.Sc (Stat Hons)
bibliography:
- bib/gravitas.bib
- bib/hakear.bib
- bib/gracsr.bib
- bib/thesisrefs.bib
link-citations: yes
---

```{r initial, echo = FALSE, cache = FALSE, results = 'hide'}
options("knitr.graphics.auto_pdf" = TRUE)
library(knitr)
opts_chunk$set(
  warning = FALSE, message = FALSE, echo = FALSE,
  fig.path = 'figure/', fig.align = 'center', fig.show = 'hold',
  cache = TRUE, external = TRUE, comment = "#>", dpi = 300,
  out.width = ifelse(is_html_output(), "100%", "\\textwidth"),
  cache.path = ifelse(is_html_output(), 'cache-gitbook/', 'cache-pdfbook/')
)
hook_output <-  knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})
```

`r if (knitr::is_latex_output()) '<!--'`

# Welcome {-}

This is the website for my PhD thesis at Monash University (Australia), titled "Visualization and analysis of probability distributions of large temporal data".

`r Sys.Date()`

`r if (knitr::is_latex_output()) '-->'`