### 2021-11-05 Rmd/06-appendix.Rmd
ensuring every chapter has #

# (APPENDIX) Appendix {-}

## Data dictionary

```{r wdi-data-dict}
```

(ref:wdi-data-dict) Data dictionary for the World Development Indicators.