### 2021-11-09 Rmd/00-c-declaration.Rmd
running after updating R to 4.1.2

---
output:
  pdf_document: default
  html_document: default
---
`r if (knitr::is_html_output()) '<!--'`

\newpage

# Declaration {-#ch:declare}

I hereby declare that this thesis contains no material which has been accepted for the award of any other degree or diploma at any university or equivalent institution and that, to the best of my knowledge and belief, this thesis contains no material previously published or written by another person, except where due reference is made in the text of the thesis.

Three publications are included in this thesis, one of which has been accepted for publication in the Journal of Computational and Graphical Statistics (@Gupta2021-gravitas). Two others have not yet been submitted. "Exploring probability distributions across cyclic temporal granularities" is the central theme of the thesis. All of the papers in the thesis were conceptualized, developed, and written by myself, the student, while working in the Department of Econometrics and Business Statistics under the supervision of Professor Rob J Hyndman and Professor Dianne Cook. When co-authors are listed, it indicates that the work was produced through active collaboration between researchers and that their contributions to team-based research have been recognized. The following table details the work, including
my and my fellow co-authors contributions:

\begin{table}
\centering\footnotesize\tabcolsep=0.12cm
\begin{tabular}{|p{1cm}|p{2cm}|p{1.5cm}|p{3.5cm}|p{3.5cm}|p{1.5cm}|}
\hline
\RaggedRight\textbf{Thesis Chapter}  & \RaggedRight\textbf{Publication Title}  & \RaggedRight\textbf{Status (published, in press, accepted or returned for revision)}  & \RaggedRight\textbf{Nature and} {\%} \RaggedRight\textbf{of student contribution} & \RaggedRight\textbf{Co-author name(s) Nature and} {\%} \RaggedRight\textbf{of Co-author’s contribution} & \RaggedRight\textbf{Co-author(s), Monash student Y/N} \\ \hline
2 & Visualizing probability distributions across bivariate cyclic temporal granularities & Published & 70\%. Concept and developing software and writing first draft & (1) Rob J Hyndman, input into manuscript 15\% (2) Dianne Cook, Concept and input into manuscript 15\% & \multicolumn{1}{c|}{N} \\
\hline
3 & Detecting distributional differences between temporal granularities
for exploratory time series analysis & To be submitted & 70\%. Concept and developing software and writing first draft & (1) Rob J Hyndman, Concept and input into manuscript 15\% (2) Dianne Cook, Concept and input into manuscript 15\%  & \multicolumn{1}{c|}{N} \\
\hline
4 & Clustering time series based on probability distributions across temporal granularities & To be submitted & 70\%. Concept and developing software and writing first draft & (1) Dianne Cook, Concept and input into manuscript 20\% (2) Rob J Hyndman, concept and input into manuscript 10\% & \multicolumn{1}{c|}{N} \\
\hline
\end{tabular}
\end{table}

I have not renumbered sections of submitted or published papers in order to generate a consistent presentation within the thesis.

\textbf{Student name:} Sayani Gupta

\textbf{Student signature:}

```{r signature, out.width = "20%", fig.align = "left", eval = FALSE}
knitr::include_graphics("img/signature.png")
```

<!-- eval = file.exists("img/signature.png") -->

\textbf{Date:} `r Sys.Date()`

`r if (knitr::is_html_output()) '-->'`