### 2022-02-02 report/reponse.Rmd
chapter 3 catherine comments done

---
title: Amendments summary
author: Sayani Gupta
output:
  MonashEBSTemplates::memo:
    includes:
      in_header: preamble2.tex
fontfamily: mathpazo
---

I thank my examiners for their thorough and constructive comments. The point by point description of changes are below: the examiners' comments are in red and my response is in black.

## Professor Catherine Hurley {-}

I appreciate all the comments and suggestions, and will incorporate them into future work. Specific comments are addressed as follows:

### Chapter 2

> Q1. In Figure 2.1 on the right hand side it should be t/24, t/(24 × 7)

> Q2. Last sentence of Section 2.3.4: fix the table reference

Done.

> Q3. In Figure 2.4, the outlier glyphs are in my opinion too big and dark and attract too much attention.
The log scale for the y-axis should be mentioned in the caption.

> Q4. The violin plot in Figure 2.5c is barely recognisable, and comparison of the distributions is challenging.
The plot deserves a bit more space.

### Chapter 3

> Q1. On page 33 there is a mention of the hakear package, without a reference.

It appears in the abstract of the paper. 
For the thesis, I have added it since the abstract acts as a summary to the chapter.

> Q2. In the Introduction state the methods are for continuous variables.

Added "for a continuous univariate dependent variable" in the last paragraph of Introduction.

> Q3. In Figure 3.1, the points are a bit hard to see, so it looks like whiskers extend beyond the data.

Changed to theme_bw() instead of default and outlier.alpha (0.5 ->0.1) and geom_jitter(alpha = 0.04 -> 0.06)

> Q4. On page 35, refer to months consistently using short or long names.

No change. Refer as months or month_year depending on the context.

> Q5. On page 35 v refers to the number of variables, in Section 3.3.3 it refers to the variable.

Changed. On p35, $v$ changed to $p$, where $p$ represents the number of variables. This is analogous to many time granularities and their pairwise combinations and $v$ represents the univariate measured variable for which distributions are constructed for different time granularities and their pairs.

> Q6. In the Figure 3.2 caption, I do not understand the comment beginning “Difference between the 90th. . . ”.
Also for the sentence “Energy consumption for (a)..”, it would clarify to insert the word “median”.

No change.
Changed to "distribution of energy usage"

> Q7. On the top of page 38, it should say “the distribution means are three standard deviations apart”.

Done.

> Q8. Maybe include a reference for gestalt theory.

To do.

> Q9. In Section 3.2.2, I find the references to the null distribution confusing. Surely that is the design in
Figure 3.3(a) only?

No change.
Null distributions refer to the case when there is no significant distributional differences between categories of one or more granularities.

> Q10. In table 3.1 Nc is the number of cyclic granularities, whereas in the text (first sentence of 3.3.2 and
3.3.3) it is m.

Add $m$ to table 3.1
$m$ = number of cyclic granularities to display together

$N_c$ refers to the total number of contextual cyclic granularities. $m$ refers to the number of cyclic granularities we are considering together in the display. For example, contextual cyclic granularities could be $hour_day$, $day_week$ and $month_year$ and we want to visualize any one granularity at a time. So $N_c = 3$ and $m=1$.

> Q11. Add v to table 3.1.

Done.
Added v:continuous univariate measured variable

> Q12. In the first sentence on the top of page 42, ordered and unordered are mixed up. Have you considered
the setting where the facet variable levels are ordered? For the within-facet ordered distances, you
could consider a distance measure that respects circular order, or choose the start level appropriate to
the display.

first part: Done.
second part: To do

> Q13. I like Figure 3.4. In (3), the dotted arcs only connect to a1 which might be misleading.

Changed.
Caption changed to incorporate that.
Within-facet distances are illustrated in Panels 3) (when categories are un-ordered, shown only with respect to a1) and Panel 4) (when categories are ordered)

> Q14. In the pairwise distance measure on page 43, is there any adjustment made for varying numbers of
observations across the levels of A and B? For example in Figure 3.3(d) if there are few values at
xlevel=1 the comparison shown is less interesting

No change.
We assume we have enough observations for each level or combination of levels to compute distributions. As long as that holds, no adjustment made for varying number of observations.

> Q15. The equations on page 44 could be tidied up. There are extra () on the definition of wpdperm and on
the residual definition at the bottom of the page. Use × not  for the equations.

Done.

> Q16. Is there a practical reason why wpdglm is not working as expected for lower nx and nf ?

To do.

> Q17. In the 3.3.1 algorithm, there is m, and then M.

Changed.
1.b) for $i \in {1, 2, \dots, M}$.

> Q18. In Table 3.2, I would suggest separating the m = 1 and m = 2 tables.

Not changed.

> Q19. The presentation of the material on the simulation study on page 48 could be improved. Where is
the notation wpdl,s used? Figure 3.5 shows the m = 2 results only. The text alludes to a simulation
involving different underlying distributions, but this is not mentioned again.

first part: How?
second part: After NQT, results are pretty similar to an underlying normal distribution.

> Q20. In Figure 3.5 the axis tick labels should be smaller. The blue and orange marks are hard to see. Maybe
show fewer panels?

To do.

> Q21. In Figure 3.6 the axis tick labels should be smaller. The blue and orange are hard to see. The caption
should refer to the rug. Maybe show fewer panels?

To do.

> Q22. In Figure 3.7 (a) the heatmaps need id labels. The grey color is missing from the legend. Maybe use a
different colour in the heatmap for the significant comparisons. State the threshold for significance in
the caption. In Table 3.3 the caption should explain the threshold. Maybe use colour instead of stars
to indicate significance?

To do.

> Q23. The link in the Acknowledgements https://github.com/Sayani07/paper-hakear is not available. Neither
are the supplementary materials. I understand this is for the paper version, I mention it for completeness.

Changed.
The link works. The repository has now been made public.

### Chapter 4

> Q1. In Section 4.1, line 5 “method of time series clustering”.

Done.

> Q2. Page 59, fourth bullet point. I found these sentences confusing.

To do.

> Q3. In the material in the bottom of page 59, make it clear from the outset you only have data on energy
use, not on property size, location, family size and so on.

To do.

> Q4. End of Chapter 4.1: incorrect reference to Section 2.7.

> Q5. I found Figure 4.1 very useful. In each box, would be helpful to put in “or”, in places where only one of
the steps listed is performed, eq Normal quantile transform or Robust scaling. As there are many steps
in the algorithm, it would be helpful to the reader to label the pipeline steps and to refer back to them
in the text. The text in the Data pre-processing box does not make sense to me.

> Q6. The section describing RS and NQT is confusing. It states RS is applied to each time series separately.
Is NQT also applied to each observation separately? How does “it could be useful to standardize it for
the selected set of significant granularities prior to computing the distances” relate to the following
bullet points?

> Q7. Page 64 “D is the Jensen-Shannon distance”.

> Q8. In Table 4.1, the R column should have 250, not 20.

> Q9. The description of the data generation on page 67 and Table 4.2 is confusing. Maybe state the
distribution of each vt.

> Q10. Where is μ in Table 4.2?

> Q11. In Section 4.3.3 what are the values of R and T?

> Q12. Figure 4.5 caption: MDS summary plots of what?

> Q13. Figure 4.7 is missing labels (a), (b) (c).

> Q14. Why do you chose to summarize 4 and 5 clusters from JS-NQT when sindex suggests 3 clusters?

> Q15. The reference to the stationarity assumption in the Discussion needs clarification.

> Q16. Page 81 “can not” should be “cannot”

> Q17. Again, the computational burden could do with more Discussion. At present, if I were to use this
method on my data, what kind of sizes are realistic to work with?